### 2016-01-22 04-evaluation.tex
Make plots fit column width + fix evaluation intro to mention subsections in correct order

This section presents an in-depth empirical analysis of the design discussed in
Section~\ref{design-section}. To this end, we assess the scalability and
efficiency of the solution using multiple criteria.
%
First, an evaluation of the \textit{strong scaling} behavior of the distributed
algorithm is presented.
%
Second, the \textit{weak scaling} of the solution is assessed with respect to increasing
numbers of targeted latent communities.
%
Next, the effects of \textit{pipelining the computation} and varying the number of latent communities on the system's
performance is analyzed.
%
Further, we evaluate the efficiency and overhead associated with \textit{network
communication} between cluster nodes.
%
Finally, we \textit{contrast} the effectiveness of scaling the computation horizontally
and vertically.

@@

\begin{figure}[t] % [htbp]
  \centering
  \epsfig{file=plots/sweep-over-K-proportional-np.eps, width=0.48\textwidth}
  \caption{Average execution time per algorithm iteration when varying
  the number of latent communities proportionally to the number of compute
  nodes.}
  \label{fig-weak-scaling}
\end{figure}

@@

\begin{figure}[t] % [htb]
  \centering
  \epsfig{file=plots/sweep-over-np-fixed-K.eps, width=0.48\textwidth}
  \caption{(a) Execution time of 2048 algorithm iterations for the
  same problem size (com-Friendster, K=1024, M=16384 n=32) across different
  cluster sizes. (b) Speedup achieved for same experiments in (a) with respect to
  8 nodes}
  \label{fig-strong-scaling}
\end{figure}

@@

%################################################
\subsection{Horizontal vs. Vertical Scalability}
\begin{figure}[t] % [htb]
  \centering
  \epsfig{file=plots/hpc-cloud.eps, width=0.48\textwidth}
  \caption{Performance comparison between the distributed implementation
  running on DAS5 and the multi-threaded solution on machine with 40 cores and
  1TB of RAM.}
  \label{fig-scale-up}
\end{figure}
One of the main drawbacks of designing a distributed solution for a given
algorithm is the inherent complexity of communication and synchronization. A
considerably simpler approach would be developing a multi-threaded version and
running it on a machine with abundant memory and CPU cores. In such a context,
access to all of the algorithm's state would be an order of magnitude faster
than RDMA. Additionally, the overhead associated with synchronizing threads is
negligible compared to using MPI primitives. To evaluate the efficacy of both
approaches we utilized SURFsara's HPC Cloud system to instantiate a virtual
machine with 40 Intel Xeon E7-4850 cores clocked at 2.00GHz cores and 1TB of
memory. The physical machine underlying
the HPC Cloud system contains 40~CPU cores and does not oversubscribe
resources. Therefore, by provisioning all 40~cores we ensured that there is no
resource contention from other users of the system. Figure~\ref{fig-scale-up}
reports the execution time per algorithm iteration for two experimental setups.
%
First, Figure~\ref{fig-scale-up}-a shows the performance of executing the
algorithm on the HPC Cloud
system with 40 and 16 cores compared to a single DAS5 node with 16~cores. This
test uses the com-DBLP data set to enable the use of a large number of
communities without running out of memory. It is clear from the results that
the performance can benefit from the additional cores provided by HPC Cloud system.
%
Further, Figure~\ref{fig-scale-up}-b tests the performance of the HPC Cloud
system compared to 64 nodes of the DAS5
cluster using the com-Friendster data set. Clearly, the parallel and
distributed implementation vastly
outperforms the single-node multi-threaded solution. Moreover, the trajectory
of both curves shows a widening gap between them suggesting that the relative
performance difference will increase for larger $K$.
%
In conclusion, the overhead of network communication in the distributed version
is more than compensated by the increasing compute power compared to
a single-node implementation.

@@

\begin{figure}[t] % [htbp]
  \centering
  \epsfig{file=plots/qperf.eps, width=0.5\textwidth}
  \caption{Performance of DKV store compared to qperf and roundtrip MPI.}
  \label{fig-qperf}
\end{figure}

>>>>

This section presents an in-depth empirical analysis of the design discussed in
Section~\ref{design-section}. To this end, we assess the scalability and
efficiency of the solution using multiple criteria.
%
First, the \textit{weak scaling} of the solution is assessed with respect to increasing
numbers of targeted latent communities.
%
Second, an evaluation of the \textit{strong scaling} behavior of the distributed
algorithm is presented.
%
Third, the effects of \textit{pipelining the computation} and varying the number of latent communities on the system's
performance are analyzed.
%
Next, we \textit{contrast} the effectiveness of scaling the computation horizontally
and vertically.
%
Further, we evaluate the efficiency and overhead associated with \textit{network
communication} between cluster nodes.
%
Finally, we apply the algorithm to large data sets and comment on its execution time.

@@

\begin{figure}[t] % [htbp]
  \centering
  \epsfig{file=plots/sweep-over-K-proportional-np.eps, width=\columnwidth}
  \caption{Average execution time per algorithm iteration when varying
  the number of latent communities proportionally to the number of compute
  nodes.}
  \label{fig-weak-scaling}
\end{figure}

@@

\begin{figure}[t] % [htb]
  \centering
  \epsfig{file=plots/sweep-over-np-fixed-K.eps, width=\columnwidth}
  \caption{(a) Execution time of 2048 algorithm iterations for the
  same problem size (com-Friendster, K=1024, M=16384 n=32) across different
  cluster sizes. (b) Speedup achieved for same experiments in (a) with respect to
  8 nodes}
  \label{fig-strong-scaling}
\end{figure}

@@

%################################################
\subsection{Horizontal vs. Vertical Scalability}
\begin{figure}[t] % [htb]
  \centering
  \epsfig{file=plots/hpc-cloud.eps, width=\columnwidth}
  \caption{Performance comparison between the distributed implementation
  running on DAS5 and the multi-threaded solution on machine with 40 cores and
  1TB of RAM.}
  \label{fig-scale-up}
\end{figure}
One of the main drawbacks of designing a distributed solution for a given
algorithm is the inherent complexity of communication and synchronization. A
considerably simpler approach would be developing a multi-threaded version and
running it on a machine with abundant memory and CPU cores. In such a context,
access to all of the algorithm's state would be an order of magnitude faster
than RDMA. Additionally, the overhead associated with synchronizing threads is
negligible compared to using MPI primitives. To evaluate the efficacy of both
approaches we utilized SURFsara's HPC Cloud system to instantiate a virtual
machine with 40 Intel Xeon E7-4850 cores clocked at 2.00GHz cores and 1TB of
memory. The physical machine underlying
the HPC Cloud system contains 40~CPU cores and does not oversubscribe
resources. Therefore, by provisioning all 40~cores we ensured that there is no
resource contention from other users of the system. Figure~\ref{fig-scale-up}
reports the execution time per algorithm iteration for two experimental setups.
%
First, Figure~\ref{fig-scale-up}-a shows the performance of executing the
algorithm on the HPC Cloud
system with 40 and 16 cores compared to a single DAS5 node with 16~cores. This
test uses the com-DBLP data set to enable the use of a large number of
communities without running out of memory. It is clear from the results that
the performance can benefit from the additional cores provided by HPC Cloud system.
%
Further, Figure~\ref{fig-scale-up}-b tests the performance of the HPC Cloud
system compared to 64 nodes of the DAS5
cluster using the com-Friendster data set. Clearly, the parallel and
distributed implementation vastly
outperforms the single-node multi-threaded solution. Moreover, the trajectory
of both curves shows a widening gap between them suggesting that the relative
performance difference will increase for larger $K$.
%
In conclusion, the overhead of network communication in the distributed version
is more than compensated by the increasing compute power compared to
a single-node implementation.

@@

\begin{figure}[t] % [htbp]
  \centering
  \epsfig{file=plots/qperf.eps, width=\columnwidth}
  \caption{Performance of DKV store compared to qperf and roundtrip MPI.}
  \label{fig-qperf}
\end{figure}
